Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-05-04 WSSSPE4_report/WSSSPE4_report.tex
minor edits to Sandra's changes

%
For talks with multiple authors, an \textsuperscript{\textasteriskcentered} indicates the author who presented the talk.
%
%\todo{group by type - idea, position, experience, demo}

The following four papers belong to the first area and tackle diverse aspects from incentives for quality software to advocating a professional society for research software to roles and degrees for research software engineers.
\begin{itemize}
\item \textbf{Michael Heroux: \emph{Idea paper: Sustainable \& Productive:
Improving Incentives for Quality Software}}~\cite{Heroux:2016ws}.
The author presents that improving quality of scientific software is a measure
to address the increased complexity in both modeling and software. To support
quality, it is essential to provide Computational Science and Engineering teams
with publication venues, funding, and  professional recognition. Broad
improvements in sustainability and productivity can be only achieved when
publishers, funding agencies and employers raise their expectations for
software quality. Additionally, software community leaders are in the unique
position to positively impact software quality by working on establishing
incentives that will spur creative and novel approaches to improve developer
productivity  and software sustainability.

@@

\item \textbf{Christopher Gwilliams: \emph{Demo: Using industrial engagement to
create and develop research ties within academia}}~\cite{Gwilliams:2016ws}.
This demo focuses on the benefits of including a  new  software engineering
degree that is influenced by, and works directly with,
industry.  Cardiff  University  introduced  a  new  industry-focused  degree,
Bachelor of Science in Applied Software Engineering (ASE) in September
2015. The engagement with industry allows students to learn through live
projects  with  real-world  applications and forge links for career opportunities.
Also, industry benefits from this course that provides an entry point for industry
to engage with the university and to develop a network of contacts between industry
and academia, which is also a benefit for the university and allows to identify
potential opportunities for new research projects and collaborators.

The second main area concerned with characteristics and needs of research software includes the following five papers.

@@

\item \textbf{Bruce Childers\textsuperscript{\textasteriskcentered}, Jack Davidson, Wayne Graves, Bernard Rous, and
David Wilkinson: \emph{Position Paper: Active Curation of Artifacts is Changing
the Way Digital Libraries will Operate}}~\cite{Childers:2016ws}.
The paper presents the implications of ``active curation systems'' that provide
executable access to
artifacts and experiments behind published results and enables
their reuse. These systems allow changing and repeating experiments to understand
how an innovation behaves in conditions beyond the ones described in a paper and
address reproducibility challenges. Active curation systems
also  enable accountability and accelerate research progress  by
giving access to complete experimental details. As these systems
take  hold,  it  is  important  to  understand  their  capabilities  and
how digital libraries should be integrated with them. The authors describe a
study underway to explore how best
to do this integration.

Three contributions present work on the third main area and elucidate about lessons learned from or visions for use cases of scientific software and communities working with and\slash or on a software package.

@@

There were 19 lightning talks presented at WSSSPE4, as follows (in order of
presentation).
%
For talks with multiple authors, an \textsuperscript{\textasteriskcentered} indicates the author who presented the talk.
%
\begin{itemize}[itemsep=1ex]
    \item \textbf{Shoaib Sufi: \emph{The Software Sustainability Institute Fellowship
    programme; supporting the social side of research software}}~\cite{Sufi:2016ws}.
    Sufi, Community Lead at the Software Sustainability Institute (SSI), discussed
    the SSI Fellowship program and the fellows, including their
    demographic breakdown and domains of practice. This program came about
    based on the institute's desire to integrate with the research community---thus
    inspiring change from within---rather than preach to it. Five SSI fellows
    attended WSSSPE4.

>>>>

%
%
%\todo{group by type - idea, position, experience, demo}

The following four papers belong to the first area, the academic environment. They address diverse aspects ranging from incentives for quality software to advocating a professional society for research software to roles and degrees for research software engineers. For talks with multiple authors, an \textsuperscript{\textasteriskcentered} indicates the author who presented the talk.

\begin{itemize}
\item \textbf{Michael Heroux: \emph{Idea paper: Sustainable \& Productive:
Improving Incentives for Quality Software}}~\cite{Heroux:2016ws}.
The author presents that improving quality of scientific software is a measure
to address the increased complexity in both modeling and software. To support
quality, it is essential to provide Computational Science and Engineering teams
with publication venues, funding, and  professional recognition. Broad
improvements in sustainability and productivity can be only achieved when
publishers, funding agencies and employers raise their expectations for
software quality. Additionally, software community leaders are in the unique
position to positively impact software quality by working on establishing
incentives that will spur creative and novel approaches to improve developer
productivity  and software sustainability.

@@

\item \textbf{Christopher Gwilliams: \emph{Demo: Using industrial engagement to
create and develop research ties within academia}}~\cite{Gwilliams:2016ws}.
This demo focuses on the benefits of including a  new  software engineering
degree that is influenced by, and works directly with,
industry.  Cardiff  University  introduced  a  new  industry-focused  degree,
Bachelor of Science in Applied Software Engineering (ASE) in September
2015. The engagement with industry allows students to learn through live
projects  with  real-world  applications and forge links for career opportunities.
Also, industry benefits from this course that provides an entry point for industry
to engage with the university and to develop a network of contacts between industry
and academia, which is also a benefit for the university and allows to identify
potential opportunities for new research projects and collaborators. \\

The second main area, concerned with characteristics and needs of research software, includes the following five papers.

@@

\item \textbf{Bruce Childers\textsuperscript{\textasteriskcentered}, Jack Davidson, Wayne Graves, Bernard Rous, and
David Wilkinson: \emph{Position Paper: Active Curation of Artifacts is Changing
the Way Digital Libraries will Operate}}~\cite{Childers:2016ws}.
The paper presents the implications of ``active curation systems'' that provide
executable access to
artifacts and experiments behind published results and enables
their reuse. These systems allow changing and repeating experiments to understand
how an innovation behaves in conditions beyond the ones described in a paper and
address reproducibility challenges. Active curation systems
also  enable accountability and accelerate research progress  by
giving access to complete experimental details. As these systems
take  hold,  it  is  important  to  understand  their  capabilities  and
how digital libraries should be integrated with them. The authors describe a
study underway to explore how best
to do this integration.\\

Three contributions focused on the third main area, elucidating lessons learned from or visions for use cases of scientific software and communities working with and\slash or on a software package.

@@

There were 19 lightning talks presented at WSSSPE4, as follows (in order of
presentation).
%
For talks with multiple authors, an \textsuperscript{\textasteriskcentered} again indicates the author who presented the talk.
%
\begin{itemize}[itemsep=1ex]
    \item \textbf{Shoaib Sufi: \emph{The Software Sustainability Institute Fellowship
    programme; supporting the social side of research software}}~\cite{Sufi:2016ws}.
    Sufi, Community Lead at the Software Sustainability Institute (SSI), discussed
    the SSI Fellowship program and the fellows, including their
    demographic breakdown and domains of practice. This program came about
    based on the institute's desire to integrate with the research community---thus
    inspiring change from within---rather than preach to it. Five SSI fellows
    attended WSSSPE4.
